---
title: "Soccer Predictions"
author: "Ashwin Madhusudan"
date: "`r Sys.Date()`"
output: 
  html_document:
    highlight: pygments
    theme: sandstone
---

<div style="margin-bottom:100px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading the Data:

```{r }
library(readr)
soccermatches <- read_csv("spi_matches.csv")
```

## Clean the Data
```{r}
soccermatchesfinal <- plyr::rename(soccermatches, c("date" = "Match_Date"))
soccermatchesfinal$Match_Date <- format(soccermatchesfinal$Match_Date, format = "%m/%d/%y")
```

## Vizualization 1:

The first exploratory visualization is a histogram indicating the number of soccer games that were played across all 8 soccer leagues in the 2016 season. As the code chunk indicates below, this was achieved by creating a subset of the original cleaned dataset specifically for 2016 season data and then created a table for this data that acts as the base for the visualization.

```{r }
soccermatchessubset2 <- subset(soccermatchesfinal,season=='2016')
mytable <- table(soccermatchessubset2$season, soccermatchessubset2$league)
barplot <- barplot(mytable, yaxp=c(0, max(mytable),5), legend = rownames(mytable), ylab = "Number of Games", col=c("light blue"), las=2, cex.names =1)
```

</div>
## Vizualization 2:

The second exploratory visualization indicates the involvement of premier leagues in the UEFA champions leagues each season from 2016-2021 in the form of a bar chart as well as a pie chart.

```{r }
soccermatchessubset3 <- subset(soccermatchesfinal,league=='UEFA Champions League')
soccermatchessubset4 <- subset(soccermatchessubset3,PremierLeagueCheck=='YES')
mytable2 <- table(soccermatchessubset4$season)
barplot(mytable2, yaxp=c(0, max(mytable2),5), xlab = "Season", ylab = "Premier League Team Involved", col=c("blue"))
pie(mytable2,main="Pie Chart of Premier League Involvment in the UCL")
```

